### 2024-02-05 README.Rmd
nb not contig

2.  As an aid to setting up contiguity structures, particularly when islands are involved, the package has a function to quickly map any contiguity structure for visual inspection. This can also be used to examine the output of `sfdep` contiguity functions. Such maps can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.

3.  If there are some contiguities present which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.

@@

*Create a contiguity list, matrix, or `sf` dataframe with a contiguity list or matrix as column "nb", while accounting for islands.*

-   **st_quickmap_contigs()**

@@

*Use the output of **st_bridges()** as both the data and contiguity inputs for a model using, for example, `mgcv`, `brms` or `inla`.*

@@

`sfdep` offers excellent tools for building contiguity structures, among other things. It has a range of functions depending on how we want to define contiguity.

@@

It also provides a number of further helper functions to examine these contiguity structures and use them in models such that the workflow is streamlined from the human side.

@@

In this French example, the `st_islands` function **st_bridges()** produces exactly the same contiguity structure as **sfdep::st_contiguity()**. This can be seen below by piping these neighbourhood outputs through the `st_islands` function **st_quickmap_contigs()** which draws an easily editable map of the applied contiguity structure. This function works with contiguities constructed from any package as long as they are in list or matrix form.

For **st_quickmap_contigs()**, the contiguities should be within an `sf` dataframe as a column called "nb". **st_bridges()** does this automatically whereas the "nb" column needs to be added when using `sfdep` functions.

@@

```{r}
g |> st_bridges("department") |> 
  st_quickmap_contigs()

@@

```{r}
g |> mutate(nb = st_contiguity(geometry)) |> 
  st_quickmap_contigs()

@@

  g |> mutate(nb = st_contiguity(geometry)) |> 
    st_quickmap_contigs() +
    geom_sf(data = g |> group_by(region) |> summarise(), 
            linewidth = 0.5, colour = "black", fill = NA) + 
    labs(title = "st_quickmap_contigs()",
         subtitle = "1. using ggplot syntax") + 
    theme_minimal() +
    theme(panel.background = element_rect(fill = "aquamarine3", color = "black"),
          axis.text = element_blank()),

  g |> mutate(nb = st_contiguity(geometry)) |> 
    st_quickmap_contigs(linkcol = "orange", 
                        bordercol = "white", 
                        pointcol = "yellow", 
                        fillcol = "black", 
                        linksize = 0.4, 
                        bordersize = 0.3, 
                        pointsize = 0.8,
                        title = "st_quickmap_contigs()",
                        subtitle = "2. using simplified arguments"),

@@

`sfdep` offers a number of different types of contiguity structure, a selection of which are shown below. These can again be conveniently visualised using the **st_quickmap_contigs()** function:

@@

g |> mutate(nb = st_geometry(g) |> 
              st_dist_band(upper = 150000)) |> 
  st_quickmap_contigs()

@@

g |> mutate(nb = st_geometry(g) |> 
              st_knn(1, symmetric = TRUE)) |> 
  st_quickmap_contigs()

@@

id <- g$code_dept
regime <- g$region
g |> 
  mutate(
    nb = st_block_nb(regime, id)
  ) |> 
  st_quickmap_contigs()

@@

g |> mutate(nb = st_contiguity(geometry) |> st_nb_lag_cumul(2)) |> 
  st_quickmap_contigs()

@@

In the context of the constituencies of England, Scotland and Wales, there will be problems due to the presence of islands which will not feature in these graphs unless we attempt to pick out these islands and set up a buffer around them. This process is, however, cumbersome and can induce contiguities which are not intended.

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   remove_islands = T)
st_quickmap_contigs(nbsf,
                    pointsize=0.05,
                    title = "st_bridges() contiguities",
                    subtitle = "no island constituencies\n(islands which remain are part of a contiguous constituency)")

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   link_islands_k = 2)
st_quickmap_contigs(nbsf,
                    pointsize=0.05, 
                    title = "st_bridges() contiguities",
                    subtitle = "islands linked to k=2 nearest constituencies")

@@

The contiguity structure which is created can be either a named list (the default) or a named matrix. Different modelling packages have different requirements for this. Furthermore, we can choose `add_to_dataframe` to be TRUE (the default) to return a dataframe with a column called `nb` which contains the named list or matrix. If FALSE, only the contiguity structure itself is returned

@@

The contiguity is here in a list form:

@@

But the "nb" column can also be a matrix, and `st_quickmap_contigs()` will still return the same map:

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   remove_islands = F,
                   link_islands_k = 2,
                   nb_structure = "matrix",
                   add_to_dataframe = T)
st_quickmap_contigs(nbsf,
                    pointsize=0.05)

@@

The matrix contiguity structure is now of the following form:

@@

There are also functions for manually changing the results of a contiguity construction. It may be the case that you want to add some additional contiguities or to remove others. For example, you may be aware from local knowledge of connectivities which are not represented by mere contiguity of polygons. The presence of tunnels or bridges across a body of water would be an example of such a situation. The functions `st_manual_join_nb()` and `st_manual_cut_nb()` do this.

@@

These manual functions can also, of course, be used to edit any of the previously discussed contiguity structures created by `sfdep`.

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |>  
  st_quickmap_contigs()

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_quickmap_contigs()

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |> 
  st_manual_join_nb(47,9) |>
  st_manual_join_nb(47,51) |>
  st_manual_join_nb(47,55) |>
  st_manual_join_nb(51,9) |>
  st_manual_join_nb(51,47) |>
  st_quickmap_contigs()

@@

Having set up a contiguity structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.

@@

An equivalent model, this time smoothing over degree_educated, can be fitted using `brms`. This package requires the contiguity to be in matrix form:

@@

Returning to the dataset used in `sfdep`, we can easily create a smooth of suicides in France in 1830 as follows. Since there are no islands, we use `st_bridges()` and it will function like `sfdep:st_contiguity()` except that it automatically adds a contiguity 'nb' column to the dataframe.

>>>>

2.  As an aid to setting up neighbourhood structures, particularly when islands are involved, the package has a function to quickly map any neighbourhood structure for visual inspection. This can also be used to examine the output of `sfdep` neighbour functions. Such maps can be used to check if the structure makes sense, given the researcher's knowledge about the geography of the study area.

3.  If there are some neighbours assigned which are not appropriate, or if you wish to add additional ones, there are functions to allow this to be done in a straightforward and openly reportable way.

@@

*Create a neighbours list, matrix, or `sf` dataframe containing a neighbours list or matrix as column "nb", while accounting for islands.*

-   **st_quickmap_nb()**

@@

*Use the output of **st_bridges()** as both the data and neighbourhood inputs for a model using, for example, `mgcv`, `brms` or `inla`.*

@@

`sfdep` offers excellent tools for building neighbourhood structures, among other things. It has a range of functions depending on how we want to define what *neighbour* should mean.

@@

It also provides a number of further helper functions to examine these neighbourhood structures and use them in models such that the workflow is streamlined from the human side.

@@

In this French example, the `st_islands` function **st_bridges()** produces exactly the same neighbourhood structure (based on contiguity) as **sfdep::st_contiguity()**. This can be seen below by piping these neighbourhood outputs through the `st_islands` function **st_quickmap_nb()** which gives a visual representation of the structure on a map. This function works with neighbourhoods constructed from any package as long as they are in list or matrix form.

For **st_quickmap_nb()**, the neighbourhoods should be within an `sf` dataframe as a column called "nb". **st_bridges()** does this automatically whereas the "nb" column needs to be added when using `sfdep` functions.

@@

```{r}
g |> st_bridges("department") |> 
  st_quickmap_nb()

@@

```{r}
g |> mutate(nb = st_contiguity(geometry)) |> 
  st_quickmap_nb()

@@

  g |> mutate(nb = st_contiguity(geometry)) |> 
    st_quickmap_nb() +
    geom_sf(data = g |> group_by(region) |> summarise(), 
            linewidth = 0.5, colour = "black", fill = NA) + 
    labs(title = "st_quickmap_nb()",
         subtitle = "1. using ggplot syntax") + 
    theme_minimal() +
    theme(panel.background = element_rect(fill = "aquamarine3", color = "black"),
          axis.text = element_blank()),

  g |> mutate(nb = st_contiguity(geometry)) |> 
    st_quickmap_nb(linkcol = "orange", 
                        bordercol = "white", 
                        pointcol = "yellow", 
                        fillcol = "black", 
                        linksize = 0.4, 
                        bordersize = 0.3, 
                        pointsize = 0.8,
                        title = "st_quickmap_nb()",
                        subtitle = "2. using simplified arguments"),

@@

`sfdep` offers a number of different types of neighbourhood structure, a selection of which are shown below. These can again be conveniently visualised using the **st_quickmap_nb()** function:

@@

g |> mutate(nb = st_geometry(g) |> 
              st_dist_band(upper = 150000)) |> 
  st_quickmap_nb()

@@

g |> mutate(nb = st_geometry(g) |> 
              st_knn(1, symmetric = TRUE)) |> 
  st_quickmap_nb()

@@

id <- g$code_dept
regime <- g$region
g |> 
  mutate(
    nb = st_block_nb(regime, id)
  ) |> 
  st_quickmap_nb()

@@

g |> mutate(nb = st_contiguity(geometry) |> st_nb_lag_cumul(2)) |> 
  st_quickmap_nb()

@@

In the context of the constituencies of England, Scotland and Wales, there will be problems due to the presence of islands which will not feature in a contiguity graph unless we attempt to pick out these islands and set up a buffer around them. This process is, however, cumbersome and can induce contiguities which are not intended.

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   remove_islands = T)
st_quickmap_nb(nbsf,
                    pointsize=0.05,
                    title = "st_bridges() contiguities",
                    subtitle = "no island constituencies\n(islands which remain are part of a contiguous constituency)")

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   link_islands_k = 2)
st_quickmap_nb(nbsf,
                    pointsize=0.05, 
                    title = "st_bridges() contiguities",
                    subtitle = "islands linked to k=2 nearest constituencies")

@@

The neighbourhood structure which is created can be either a named list (the default) or a named matrix. Different modelling packages have different requirements for how this information should be presented. Furthermore, we can choose `add_to_dataframe` to be TRUE (the default) to return a dataframe with a column called `nb` which contains the named list or matrix. If FALSE, only the neighbourhood structure itself is returned

@@

The neighbourhoods are here in a list form:

@@

But the "nb" column can also be a matrix, and `st_quickmap_nb()` will still return the same map:

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   remove_islands = F,
                   link_islands_k = 2,
                   nb_structure = "matrix",
                   add_to_dataframe = T)
st_quickmap_nb(nbsf,
                    pointsize=0.05)

@@

The matrix neighbourhood structure is now of the following form:

@@

There are also functions for manually changing the results of a neighbourhood construction. It may be the case that you want to add some additional links or to remove others. For example, you may be aware from local knowledge of connectivities which are not represented by mere contiguity of polygons. The presence of tunnels or bridges across a body of water would be an example of such a situation. The functions `st_manual_join_nb()` and `st_manual_cut_nb()` do this.

@@

These manual functions can also, of course, be used to edit any of the previously discussed neighbourhood structures created by `sfdep`.

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |>  
  st_quickmap_nb()

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_quickmap_nb()

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |> 
  st_manual_join_nb(47,9) |>
  st_manual_join_nb(47,51) |>
  st_manual_join_nb(47,55) |>
  st_manual_join_nb(51,9) |>
  st_manual_join_nb(51,47) |>
  st_quickmap_nb()

@@

Having set up a neighbourhood structure and embedded it as a named list or matrix within the original `sf` dataset as column `nb`, there are some functions to make it easy to quickly perform ICAR smoothing, augment the original dataframe with these predictions, and visualise them.

@@

An equivalent model, this time smoothing over degree_educated, can be fitted using `brms`. This package requires the neighbourhoods to be in matrix form:

@@

Returning to the dataset used in `sfdep`, we can easily create a smooth of suicides in France in 1830 as follows. Since there are no islands, we use `st_bridges()` and it will function like `sfdep:st_contiguity()` except that it automatically adds a neighbourhood 'nb' column to the dataframe.
